---
title: "Growth Portfolio"
author: "Dylan Projansky"
date: "10/28/2017"
output: word_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Limpet growth was tracked on a combination of over 200 different plates for a span of 6 months in 2014-2015. The time frame began in June of 2014 and concluded in January of 2015. Limpets were identified by using a tag color and ID. Limpets were tracked by their Area in mm2. The difference between months allowed for the calculation of the growth in mm2. The below functions plot: individual plate growth between months, three plates worth of growth between months, limpet growth on a single plate throughout the six months, and the creation of a data frame in a long frame for an individual plate throughout the six months. 

``` {r importfulldata, echo= TRUE, warning= FALSE}
# Prepare library for reading xlsx file 
library(readxl)

# df= read_xlsx(file.choose())
df= read_xlsx("HMS_2014_limpet_growth_master_20171018.xlsx")
```

``` {r preparelimpet, echo= TRUE, warning= FALSE}
# Combine the Tag color and Tag number for each month 
df$Tag.color = paste(df$Tag.color, df$Tag.number.Jun, sep= '')
df$Tag.color__1 = paste(df$Tag.color__1, df$Tag.number.Jul, sep= '')
df$Tag.color__2= paste(df$Tag.color__2, df$Tag.number.Aug, sep = '')
df$Tag.color__3= paste(df$Tag.color__3, df$Tag.number.Sep, sep= '')
df$Tag.color__4= paste(df$Tag.color__4, df$Tag.number.Oct, sep= '')
df$Tag.color__5= paste(df$Tag.color__5, df$Tag.number.Nov, sep= '')
df$Tag.color__6= paste(df$Tag.color__6, df$Tag.number.Dec, sep= '')
df$Tag.color__7= paste(df$Tag.color__7, df$Tag.number.Jan, sep = '')

limpet= subset(df, select = c (2:5, 9, 16:18, 21, 24, 31, 34:36, 39, 42, 49,                              52:54, 57, 60, 67, 70:72, 75, 78, 85, 88:90, 93, 
                            96, 103, 106:108, 111, 114, 121, 124:126, 129, 
                            132, 139)) 

# Rename the Tag headers from Tag.color to Tag.Month 
colnames(limpet)[4]= 'Tag.Jun'
colnames(limpet)[8]= 'Tag.Jul'
colnames(limpet)[14]= 'Tag.Aug'
colnames(limpet)[20]= 'Tag.Sep'
colnames(limpet)[26]= 'Tag.Oct'
colnames(limpet)[32]= 'Tag.Nov'
colnames(limpet)[38]= 'Tag.Dec'
colnames(limpet)[44]= 'Tag.Jan'

# Convert all of the Areas into numeric values capable of determining the difference between them 
# Ignore the NAs introduced by coercion 
limpet$AreaJun.mm2= as.numeric(limpet$AreaJun.mm2)
limpet$AreaJul.mm2= as.numeric(limpet$AreaJul.mm2)
limpet$AreaAug.mm2= as.numeric(limpet$AreaAug.mm2)
limpet$AreaSep.mm2= as.numeric(limpet$AreaSep.mm2)
limpet$AreaOct.mm2= as.numeric(limpet$AreaOct.mm2)
limpet$AreaNov.mm2= as.numeric(limpet$AreaNov.mm2)
limpet$AreaDec.mm2= as.numeric(limpet$AreaDec.mm2)
limpet$AreaJan.mm2= as.numeric(limpet$AreaJan.mm2)

# Convert the dates to month-year format 
limpet$Date= as.Date.character(limpet$Date, "%Y%m%d" )
limpet$Date2= as.Date.character(limpet$Date2, "%Y%m%d" )
limpet$Date3= as.Date.character(limpet$Date3, "%Y%m%d" )
limpet$Date4= as.Date.character(limpet$Date4, "%Y%m%d" )
limpet$Date5= as.Date.character(limpet$Date5, "%Y%m%d" )
limpet$Date6= as.Date.character(limpet$Date6, "%Y%m%d" )
limpet$Date7= as.Date.character(limpet$Date7, "%Y%m%d" )
limpet$Date8= as.Date.character(limpet$Date8, "%Y%m%d" )

```

## 1 Plate Function 
The below function creates a plot of the difference between two months, and plots the tag color and number next to the point. Lines are added at the growth mark of ±5mm2. This number was chosen as an estimate of an acceptable amount of area to be gained or lost. Growth within these red lines is assumed to be accepted. 
``` {r limpetfunction, echo= TRUE}
one.plate= function(plate, month1, month2){
  # Subset data for only the single plate
  temp= as.data.frame(limpet[limpet$Plate == plate,])
  
  # Pick out the first month Area column to be looked at
  startMonth= paste('Area', month1, '.mm2', sep= '' )
  
  # Pick out the second month Area column to be looked at 
  endMonth= paste('Area', month2, '.mm2', sep = '')
  
  # Look at the tag ID for each limpet 
  tagMonth= paste('Tag.', month1, sep = '')
  
  # Look at the replacement column 
  replacementMonth= paste('Replacement', month2, sep = '')
  
  # Create a column that calculates the difference between areas 
  temp$diff= temp[, endMonth] -  temp[, startMonth]
  
  # If replacement is TRUE, then we will make the difference be NA 
  as.character(temp[, replacementMonth])
  temp$diff[temp[,replacementMonth] == TRUE]= NA
  
  temp= temp[!is.na(temp$diff),]
  
  # Calculate the max/min difference values 
  maxdiff= max(temp$diff, na.rm = TRUE)
  mindiff= min(temp$diff, na.rm= TRUE) 
  ylims= c(mindiff, maxdiff)
  
  if(ylims[1]< -6){
    # do nothing
  } else {
      ylims[1] = -6
      }
  
  if(ylims[2]> 6){
    # do nothing
  } else {
      ylims[2] = 6
      }
  
  plot(x= temp$Plate, y= temp$diff, 
        xlim = c(plate - 2, plate + 2), ylim= ylims,
        xlab= "Plate Number", 
        ylab= expression(paste("Growth mm" ^2, sep= '')), 
        main= paste("Growth of Plate", plate, "from", month1, "to", month2) )
  text(x= temp$Plate, y=temp$diff, 
       labels = temp[, tagMonth],
       cex=0.7, pos = 4)
  abline(b= 0, h= 5, col= "red")
  abline(b=0, h= -5, col= "red")
  abline(b=0, h=0, col= "black", lty=2) 
                                  
}
``` 
# Working Plot Anaylsis 
  The below plots were created to fact-check the data set. Plates which included limpets outside of the growth of ±5 mm2 were set aside to look at the pictures between months. 
  Many of the growths that fall outside of this 5 mm2 range may be explained by a variation in the calibration of the picture measurements. However, some of the limpets may have been partially covered during one month, or incorrectly identified. The below set examines the plates up to 20 as a working example. 

## Plate 1: 
  * W5 is partially covered by another limpet in September
  * W5 might have been replaced from October to November
  * Coloration in pictures look slightly different
``` {r plate1_W5, echo= TRUE}
one.plate(1, "Aug", "Sep")
one.plate(1, "Sep", "Oct")
one.plate(1, "Oct", "Nov")

```
  
## Plate 2:  
The limpets of plate 2 all maintained a growth of ±5 mm2 per month 

## Plate 3: 
  * Or68 growth: Nov (62.83), Dec(86.67), Jan(62.94)
  * Believe that the measurement in December is incorrect
  * Change in growth from Nov to December is 0.11 
``` {r plate3_Or68, echo= TRUE}
one.plate(3, "Nov", "Dec")
one.plate(3, "Dec", "Jan")

```
  
## Plate 7: 
The limpets of plate 7 all maintained a growth of ±5 mm2 per month 

## Plate 8: 
  * W19 increases by about 7 
  * Change seems to be due to bag angle in Jun of W19 
  
  * Or47 increases by about 6 
  * This growth may be accurate
``` {r plate8, echo= TRUE}
one.plate(8, "Jun", "Jul")
one.plate(8, "Jul", "Aug")

```

## Plate 9: 
The limpets of plate 9 all maintained a growth of ±5 mm2 per month 

## Plate 10: 
  * B78 looks like the same limpet and may have grown an accurate ammount from the months of August to October 
  * Or22 becomes Or56. Looks like the same limpet with new tag, but not 100%
  * W32 also becomes W63 from Nov to Dec 
``` {r plate10, echo= TRUE}
one.plate(10, "Aug", "Sep")
one.plate(10, "Sep", "Oct")
one.plate(10, "Nov", "Dec")

```

## Plate 11 
  * Gr11 seems to be accurate and could be due to the camera angle
  * I believe W33 and W62 are not the same limpet 
  * W62 and Or62 both have a difference of ±40
  * This is because we flipped the W62 and Or62 placement in the Excell file
``` {r plate11, echo= TRUE}
one.plate(11, "Jun", "Jul")
one.plate(11, "Sep", "Oct")

```

## Plate 12 
  June to July 
    * Or19 and B64 may both fall outside of the ±5 range because of the camera angle 
  August to September 
    * Gr7 may also be due to the camera angle 
  October to Novmber 
    * W87 is partially covered in October and then is alone in November 
``` {r plate12, echo= TRUE}
one.plate(12, "Jun", "Jul")
one.plate(12, "Aug", "Sep")
one.plate(12, "Oct", "Nov")

```
    
## Plate 16:
  * Or28, Or26, Gr57: All appear accurate with some variance to camera angle and overlap 
  * B88: overlap in September  
  * Or16: overlap in December 
``` {r plate16, echo= TRUE}
one.plate(16, "Jul", "Aug")
one.plate(16, "Aug", "Sep")
one.plate(16, "Dec", "Jan")
 
```

## Plate 17: 
  *Or27: heavily chipped in November, seems accurate 
``` {r plate17, echo= TRUE}
one.plate(17, "Nov", "Dec")

``` 
  
## Three plate function 
The function below is a direct addition to the previous function which plots one plate growth between two consecutive months. This function plots three plates' growth between consectuive months. The function is capable of plotting any plates, but graphs best when the plate numbers are close together, example: 7, 8, 9. Like above, the months are three letter abbreviations and are written in quotation marks, example: July = "Jul", November = "Nov". 

``` {r 3platefunction, echo= TRUE} 
three.plate= function(plate1, plate2, plate3, month1, month2){
  
  temp1= as.data.frame(limpet[limpet$Plate == plate1,])
  temp2= as.data.frame(limpet[limpet$Plate == plate2,])
  temp3= as.data.frame(limpet[limpet$Plate == plate3,])
  temp= rbind(temp1, temp2, temp3)
  
  startMonth= paste('Area', month1, '.mm2', sep= '' )
  
  endMonth= paste('Area', month2, '.mm2', sep = '')
  
  tagMonth= paste('Tag.', month1, sep = '')
  
  replacementMonth= paste('Replacement', month2, sep = '')
  
  temp$diff= temp[, endMonth] -  temp[, startMonth]
  
  as.character(temp[, replacementMonth])
  
  temp$diff[temp[,replacementMonth] == TRUE]= NA
  
  temp= temp[!is.na(temp$diff),]
 
  # Take into account the total max/min for all three plates  
  ymax= max(temp$diff, na.rm= TRUE)
  ymin= min(temp$diff, na.rm= TRUE) 
  ylims= c(ymin, ymax)
  
  if(ylims[1]< -6){
    # do nothing
  } else {
      ylims[1] = -6
      }
  
  if(ylims[2]> 6){
    # do nothing
  } else {
      ylims[2] = 6
      }
  
  # Plot the first plate only taking into account max/min 
  plot(x= temp$Plate[temp$Plate == plate1], 
       y= temp$diff[temp$Plate == plate1], 
       col= "blue", pch= 1, 
       xlim = c(plate1 -3 , plate3 + 3), 
       ylim= ylims,
       xlab= "Plate Number", 
       ylab= expression(paste("Limpet Growth mm" ^2, sep= '')),
       cex.lab= 0.9, 
       main= paste("Limpet Growth from", month1, "to", month2) )
  # Add points for the second plate 
  points(x= temp$Plate[temp$Plate == plate2], 
         y= temp$diff[temp$Plate == plate2], 
         col="red", pch= 2 )
  # Add points for the third plate 
  points(x= temp$Plate[temp$Plate == plate3], 
         y= temp$diff[temp$Plate == plate3], 
         col= "green",pch= 3 )
  text(x= temp$Plate, y=temp$diff, 
       labels = temp[, tagMonth],
       cex=0.7, pos = 4)
  abline(b= 0, h= 5, col= "red")
  abline(b=0, h= -5, col= "red")
  abline(b=0, h=0, col= "black", lty=2) 
  # Create a legend which identifies the plate with the color 
  legend("topright", legend= c(plate1, plate2, plate3), 
         col = c("blue", "red", "green"), 
         pch = c(1, 2, 3), title = "Plate Number", cex = 0.75 ) 
                                  
}

``` 

## Plate 18:
  * The large growth of plate 18 limpets may be due to human errors of camera angle or calculating the mm2, but the limpet measurements appear reasonalbe 

## Plate 19: 
  * W47 falls outside of the boundaries from August to September most likely do to human error in the measurements 

## Plate 20: 
  August to September 
    * Or34: September measurements may be low 
  September to October 
    * Or82 is chipped in October 
  December to January 
    * W57 appears to be accurate in the pictures 
    
 Below is an example of the Three Plate Tracker Function for the month of August to September    
```{r threeplot, echo= TRUE}
three.plate(18, 19, 20, "Aug", "Sep")

```

``` {r five_plate_growth, echo= FALSE}
## Five Plate Function 
## Similar to the one plate function and the three plate function 
## to plot the difference in limpet growth, this function does the same
## with five total plates. The function is very similar to the three
## plate with the addition of two more plates being plotted. A legend is
## created the same way and the function is best run when 5 plates are in
## a row (1, 2, 3, 4, 5). 

five.plate= function(plate1, plate2, plate3, plate4, plate5, month1, month2){
  
  temp1= as.data.frame(limpet[limpet$Plate == plate1,])
  temp2= as.data.frame(limpet[limpet$Plate == plate2,])
  temp3= as.data.frame(limpet[limpet$Plate == plate3,])
  temp4= as.data.frame(limpet[limpet$Plate == plate4,])
  temp5= as.data.frame(limpet[limpet$Plate == plate5,])
  temp= rbind(temp1, temp2, temp3, temp4, temp5)
  
  startMonth= paste('Area', month1, '.mm2', sep= '' )
  
  endMonth= paste('Area', month2, '.mm2', sep = '')
  
  tagMonth= paste('Tag.', month1, sep = '')
  
  replacementMonth= paste('Replacement', month2, sep = '')
  
  temp$diff= temp[, endMonth] -  temp[, startMonth]
  
  as.character(temp[, replacementMonth])
  
  temp$diff[temp[,replacementMonth] == TRUE]= NA
  
  temp= temp[!is.na(temp$diff),]
  
  ymax= max(temp$diff, na.rm= TRUE)
  ymin= min(temp$diff, na.rm= TRUE) 
  
  plot(x= temp$Plate[temp$Plate == plate1], 
       y= temp$diff[temp$Plate == plate1], 
       col= "blue", pch= 1, 
       xlim = c(plate1 -3 , plate3 + 3), 
       ylim= c(ymin -2, ymax +2),
       xlab= "Plate Number", ylab= "Limpet Growth mm2", 
       main= paste("Limpet Growth from", month1, "to", month2) )
  points(x= temp$Plate[temp$Plate == plate2], 
         y= temp$diff[temp$Plate == plate2], 
         col="red", pch= 2 )
  points(x= temp$Plate[temp$Plate == plate3], 
         y= temp$diff[temp$Plate == plate3], 
         col= "green",pch= 3 )
  points(x= temp$Plate[temp$Plate == plate4], 
         y= temp$diff[temp$Plate == plate4], 
         col= "orange",pch= 4 )
  points(x= temp$Plate[temp$Plate == plate5], 
         y= temp$diff[temp$Plate == plate5], 
         col= "black",pch= 5 )
  text(x= temp$Plate, y=temp$diff, 
       labels = temp[, tagMonth],
       cex=0.7, pos = 4)
  abline(b= 0, h= 5, col= "red")
  abline(b=0, h= -5, col= "red")
  legend("topleft", legend= c(plate1, plate2, plate3, plate4, plate5), 
         col = c("blue", "red", "green", "orange", "black"), 
         pch = c(1, 2, 3, 4, 5), title = "Plate Number", cex = 0.75 ) 
                                  
}

``` 

## Tracker Function 
The below function is designed to track the growth of a single plate throughout the entire six month time frame. Limpets are tracked by a line connecting a single, same limpet in time. Lines that end or do not start until later in the plot show a limpet that was either remarked, new, or left. Some tags in the final month have a color (G) and then "no tag". This is because this was the last month of tracking, and limpets that lost a tag but appeared to be the same from the previous month were given this tag. 
``` {r dataframe_tracker, echo= TRUE, warning= FALSE}
tracker= function(plate1){
  # Subset the full data frame to only provide for a single plate 
  data= as.data.frame(limpet[limpet$Plate == plate1,])
  # Only include the columns for Plate, Area, Growth, Replacement, and Tag
  data= subset(data, select= c(1:6, 8:12, 14:18, 20:24, 26:30, 32:36, 38:42,
                               44:47) )

  # Subset out each month's data into a single data frame 
  # Add a column for Number which tracks a single limpet throughout 
  # each month. 
  # Number 1 should be the same limpet in all months 
  sub1= subset(data, select = c("Date2", "Tag.Jul", "ReplacementJul", "AreaJul.mm2", "GrowthJul.Area.mm2"))
  colnames(sub1)= c("Date", "Tag", "Replacement", "Area", "Growth")
  sub1$Num= 1:nrow(data)
  sub2= subset(data, select= c("Date3", "Tag.Aug", "ReplacementAug", "AreaAug.mm2", "GrowthAug.Area.mm2"))
  colnames(sub2)= c("Date", "Tag", "Replacement", "Area", "Growth")
  sub2$Num= 1:nrow(data)
  sub3= subset(data, select = c("Date4", "Tag.Sep", "ReplacementSep", "AreaSep.mm2","GrowthSep.Area.mm2"))
  colnames(sub3)= c("Date", "Tag", "Replacement", "Area", "Growth")
  sub3$Num= 1:nrow(data)
  sub4= subset(data, select = c("Date5", "Tag.Oct", "ReplacementOct", "AreaOct.mm2","GrowthOct.Area.mm2"))
  colnames(sub4)= c("Date", "Tag", "Replacement", "Area", "Growth")
  sub4$Num= 1:nrow(data)
  sub5= subset(data, select = c("Date6", "Tag.Nov", "ReplacementNov", "AreaNov.mm2","GrowthNov.Area.mm2"))
  colnames(sub5)= c("Date", "Tag", "Replacement", "Area", "Growth")
  sub5$Num= 1:nrow(data)
  sub6= subset(data, select = c("Date7", "Tag.Dec", "ReplacementDec", "AreaDec.mm2","GrowthDec.Area.mm2"))
  colnames(sub6)= c("Date", "Tag", "Replacement", "Area", "Growth")
  sub6$Num= 1:nrow(data)
  sub7= subset(data, select = c("Date8", "Tag.Jan", "ReplacementJan", "AreaJan.mm2","GrowthJan.Area.mm2"))
  colnames(sub7)= c("Date", "Tag", "Replacement", "Area", "Growth")
  sub7$Num= 1:nrow(data)
  
  # Combine all of the data frames into a single long-form data frame 
  longdata= rbind(sub1, sub2, sub3, sub4, sub5, sub6, sub7)
  
  # Clean-up 
  rm(sub1, sub2, sub3, sub4, sub5, sub6, sub7)
  
  # Convert the Growth column to numeric
  longdata$Growth= as.numeric(longdata$Growth)
  
  for(i in 1:length(unique(longdata$Num)) ) {
   # Go through the data frame. If the Number column is 1, then make a plot
    # Plot the Growth on the y-axis, Date on the x-axis 
    # Include the TagID for the limpet throughout the plot 
    if(i == 1){
      plot(x= longdata$Date[longdata$Num == i], 
           y= longdata$Growth[longdata$Num == i], col= i, 
           xlab= "Months", 
           ylab= expression(paste("Growth mm" ^2, sep= '')),
           ylim= c(min(longdata$Growth, na.rm= TRUE),
                   max(longdata$Growth, na.rm = TRUE) ),
           xlim = c(min(longdata$Date), as.Date('2015-02-01')), 
           main= paste("Six Month Tracker of Plate", plate1) ) 
      text(x= longdata$Date[longdata$Num == i], 
           y= longdata$Growth[longdata$Num == i], 
         labels = longdata[longdata$Num == i, "Tag"],
         cex=0.7, pos = 4)
      # If Number ≠ 1, then make a points for each Number (2-etc.)
    } else {
      points(x= longdata$Date[longdata$Num == i], 
             y= longdata$Growth[longdata$Num == i], col=i+1)
      text(x= longdata$Date[longdata$Num == i], 
           y= longdata$Growth[longdata$Num == i], 
         labels = longdata[longdata$Num == i, "Tag"],
         cex=0.7, pos = 4)
  }
  }  
  
  
for(i in 1:length(unique(longdata$Num)) ) {
    # For those limpets, if the number is the same, then connect them
    # with a line 
    if(i == 1){  
      lines(x= longdata$Date[longdata$Num == i], 
           y= longdata$Growth[longdata$Num == i], col= i)
    } else {
      lines(x= longdata$Date[longdata$Num == i], 
             y= longdata$Growth[longdata$Num == i], col=i+1)
    }
}
  abline(b=0, h=0, col= "black", lty=2) 
}
```

Below is an example of the tracker function on plate 21. 
```{r tracker, echo= TRUE, warning= FALSE }
tracker(21)
```

## Generate Dataframe Funciton 
The below function takes the entire original dataframe and converts it into a long version which includes the Date, Tag, Replacement, Area.mm2, Growth.mm2, and also a column for Number. The number column is designed to track a single limpet through time.  
```{r save_df, echo= TRUE}
dataframe= function(plate1){
  data= as.data.frame(limpet[limpet$Plate == plate1,])
  data= subset(data, select= c(1:6, 8:12, 14:18, 20:24, 26:30, 32:36, 38:42,
                               44:47) )

  sub1= subset(data, select = c("Date2", "Tag.Jul", "ReplacementJul", "AreaJul.mm2", "GrowthJul.Area.mm2"))
  colnames(sub1)= c("Date", "Tag", "Replacement", "Area", "Growth")
  sub1$Num= 1:nrow(data)
  sub2= subset(data, select= c("Date3", "Tag.Aug", "ReplacementAug", "AreaAug.mm2", "GrowthAug.Area.mm2"))
  colnames(sub2)= c("Date", "Tag", "Replacement", "Area", "Growth")
  sub2$Num= 1:nrow(data)
  sub3= subset(data, select = c("Date3", "Tag.Sep", "ReplacementSep", "AreaSep.mm2","GrowthSep.Area.mm2"))
  colnames(sub3)= c("Date", "Tag", "Replacement", "Area", "Growth")
  sub3$Num= 1:nrow(data)
  sub4= subset(data, select = c("Date4", "Tag.Oct", "ReplacementOct", "AreaOct.mm2","GrowthOct.Area.mm2"))
  colnames(sub4)= c("Date", "Tag", "Replacement", "Area", "Growth")
  sub4$Num= 1:nrow(data)
  sub5= subset(data, select = c("Date5", "Tag.Nov", "ReplacementNov", "AreaNov.mm2","GrowthNov.Area.mm2"))
  colnames(sub5)= c("Date", "Tag", "Replacement", "Area", "Growth")
  sub5$Num= 1:nrow(data)
  sub6= subset(data, select = c("Date6", "Tag.Dec", "ReplacementDec", "AreaDec.mm2","GrowthDec.Area.mm2"))
  colnames(sub6)= c("Date", "Tag", "Replacement", "Area", "Growth")
  sub6$Num= 1:nrow(data)
  sub7= subset(data, select = c("Date7", "Tag.Jan", "ReplacementJan", "AreaJan.mm2","GrowthJan.Area.mm2"))
  colnames(sub7)= c("Date", "Tag", "Replacement", "Area", "Growth")
  sub7$Num= 1:nrow(data)
  
  longdata= rbind(sub1, sub2, sub3, sub4, sub5, sub6, sub7)
  
  # Clean-up 
  rm(sub1, sub2, sub3, sub4, sub5, sub6, sub7)
  
  longdata$Growth= as.numeric(longdata$Growth)
  
  filename= paste("limpet_growth_plate", plate1, ".csv", sep = '')
  write.csv(longdata, file = filename )

}
``` 