---
title: "Week6"
author: "Dylan Projansky"
date: "9/27/2017"
output: word_document
---

``` {r functionExample1}
myfunction= function(x) {
  print(x)
}
```
The above code creates a new function 'myfunction'. It takes a single argument 'x' and when the function executes, it will print out the contents of 'x' 

``` {r functionExample2}
myfunction2= function(x, y) {
  z= x + y 
  z
}

result= myfunction2 (2, 4)
print(result)
```
Note that the code above, we run the function and assign the results to a new variable called 'result'. To actually show what was calculated, we have to print out 'result'. If we ran the function without assigning its results to a variable, the contents of 'z' to the console. 

``` {r SEfunction}
mySE= function(x) { 
  # Assume the input x is a vector of numbers 
  # First calculate the variance 
  myvar= var(x)
  # Then calculate the number of values in the sample 
  myn= length(x)
  # Next calculate the square root of the variance, divide by the sample size
  mySE= sqrt(myvar/myn)
  # The value returned will be the last object created in the funciton, in this case my SE
  mySE 
  # Repeating just the last variable here will cause the result to print out when run 
}

``` 

Now we can test the function 
``` {r test mySE}
x= c(1, 1, 3, 3, 6, 6)
mySE(x) # Result should be 0.9189366

```

There are functions in R that will sometimes only take the name of another function as an argument. One example we've used is the 'aggregate' function, which we used to calculate means or standard deviations of different groups in a data frame. If you write your own custom function, you can supply it to 'aggregate' as an argument and have it carry out your custom function on each group. 

``` {r aggregaeCustomFunction}
# Make a practice data frame 
mydf= data.frame(Group= c(rep('A', 6), rep('B', 6)),
                 Values= c(1, 2, 3, 4, 5, 6, 11, 22, 33, 44, 55, 66))
mydf # Print it out 
# Calculate the SE of group A (rows 1-6) 
mySE(mydf$Values[1:6])
# Calculate the SE of group B (rows 7-12) 
mySE(mydf$Values[7:12]) 

# Now use our custom mySE function inside aggregate() to accomplish the same thing 
aggregate(Values~Group, data= mydf, FUN= mySE) 

``` 
