---
title: "Week8_homework"
author: "Dylan Projansky"
date: "10/11/2017"
output: word_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

``` {r importingFiles, cachee= TRUE}

fnames= dir("../data/temperature_trials", full.names = TRUE)

for(i in 1:length(fnames) ) {
  # open each file in turn 
  temp= read.csv(fnames[i])
  
  loc= regexpr(pattern= "Trial[0-9]*", text= fnames[i])

  trialNumber= as.numeric(substr(fnames[i], start= loc[[1]][1] + 5, 
         stop= loc[[1]][1] + attr(loc, "match.length") -1)) 

  temp$TrialNumber= trialNumber
  
  temp$DateTime= as.POSIXct(temp$DateTime, origin= '1970-1-1', tz= "")
  difftimes= difftime(temp$DateTime, temp$DateTime[1], units = "mins")
  # Create column of data for elapsed time in minutes 
  
  temp$DiffTimes = difftimes
  temp$ElaspeTimeMin= as.numeric(temp$DiffTimes)
  
  if(i == 1){
    res= temp
    # For the first dataset in fnames, create a plot 
    plot(x= res$ElaspeTimeMin, y= res$TC0, type = "l", xlim = c(0, 180), ylim = c(15, 35), 
         main = "Channel TC0 for each Trial", xlab = "Elapsed Time, minutes", 
         ylab= "Temperature, C", las= 1) 
  } else {
    # For all additional datasets in fnames, add a new line to the existing plot 
    # For each of those points, make the new plot in a new color. 
    points(x= temp$ElaspeTimeMin, y= temp$TC0, type = "l", col= i)
    res= rbind(res, temp)
  }

}

```


This first plot was created by using the initial loop for inputting the data. Therefore, should more temperature trials be added, the plot will be updated continuously. This is done by taking the first trial [1] and creating a singular plot with it. If i does not equal 1, then that trial will be added using the points function and will change the color depending on the value for i. 

