---
title: "Week11"
author: "Dylan Projansky"
date: "11/1/2017"
output: 
  word_document: 
    fig_caption: true 
---

```{r setup, include=TRUE, echo= TRUE}
knitr::opts_chunk$set(echo = FALSE, fig.path = 'figs/', cache= TRUE)
```

This example will be working with the built-in 'pressure' data" 
```{r pressureData, echo=TREU}
summary(pressure) 
pressure= pressure
```

## Increasing the size of the image 
Adding a 'dpi=xxx' argument inside the chunk header will create a larger png file 
{r pressure, dpi= 300} will increase the file dimensions to roughly 1500x1200px (5"x 300dpi = 1500px)

``` {r pressure2, dpi= 300}
plot(pressure, col= 'blue')
```

## Specifyng the image dimensions explicityly 
Adding the arguments for 'fig.width' and 'fig.height' inside the R markdown chunk

``` {r pressure3, dpi= 300, fig.width=4, fig.height=7}
plot(pressure, col= 'red')
```

## Saving images as vector format files (pdf) 
{r pressure4, dpi=300, fig.width=7, fig.height=5, dev=c('png', 'pdf')}
Knit both a png version in the word document and a pdf version 
``` {r pressure4, dpi=300, fig.width=7, fig.height=5, dev=c('png', 'pdf')}
plot(pressure) 
```

## ggplot2 tutorial 

``` {r ggplot2} 
library(ggplot2)
diamonds= diamonds # make diamonds a data fram visible in the workspace 

ggplot(diamonds) # if onyl the dataset is known 
ggplot(diamonds, aes(x= carat) ) # if only X-axis is known 
# the Y-axis can be specified in respective geoms 
ggplot(diamonds, aes(x= carat, y= price)) # if both X and Y axes are fixed for all layers

ggplot(diamonds, aes(x= carat, y= price)) + geom_point()
# This will plot a scatter plot 

ggplot(diamonds, aes(x= carat, y= price, color= cut)) + geom_point()

ggplot(diamonds, aes(x= carat, y= price, color= cut)) + geom_point()+ geom_smooth()

ggplot(diamonds, aes(x=carat, y=price)) + geom_point(aes(color=cut)) + geom_smooth()
# Line only goes through the aggregate of carat and price and not individual cuts 

ggplot(diamonds, aes(x=carat, y=price, color=cut, shape=color)) + geom_point()

ggplot(diamonds, aes(x=carat, y=price)) + geom_point(aes(color=cut), alpha= 0.3) + geom_smooth()

ggplot(diamonds, aes(x= carat, y= price, color= cut)) + geom_point()+ labs(title= "Scatterplot", x= "Carat", y= "Price")

gg= ggplot(diamonds, aes(x= carat, y= price, color= cut)) + geom_point()

gg= gg + labs(title= "Scatterplot", x= "Carat", y= "Price")
print(gg)


gg1= gg+ theme(plot.title = element_text(size=30, face= "bold"), 
               axis.text.x = element_text(size=18), 
               axis.text.y = element_text(size=18), 
               axis.title.x = element_text(size=25), 
               axis.title.y = element_text(size=25)) + 
  scale_color_discrete(name= "Cut of diamonds") 
print(gg1)

gg+ facet_wrap (~cut, ncol = 3)

``` 

